---
title: "harvest"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sensitivity)
library(tidyverse)
library(deSolve)
library(here)
```

# Proportional harvest rate 

Generate and run a model of harvest
- note the model cannot start at zero because then there is nothing to grow 
- note that we are taking a proportional harvest 
```{r harvest}
source(here("R", "dharvest.R"))
dharvest
```


```{r harvest}
# try it out
tm = seq(from=1, to=500) #time 
Pinitial = 1 #start with a sapling of one 
gps = list(harv=0.18, K=100, r=0.2) #parameters 

res = ode(Pinitial,tm, dharvest, gps) #put everything into the ODE solver 
colnames(res)=c("time","carbon")

ggplot(as.data.frame(res), aes(time, carbon))+geom_point()
```


```{r harvest}
# what if we harvest a a much greater rates

# lets vary the harvest rates from 0.0 to 0.4
harvestr = seq(from=0.0, to=0.4, by=0.025)

#we want to run our model using all these different harvest rates 

# save all of the trajectories
```


```{r harvest}
# use a wrapper function to just return the carbon trajectories
#the wrapper will run the ODE for each harvest rate 

getcarbon = function(Pinitial, tm, harv, K, r, hfunc) {
  gps = list(harv=harv, K=K, r=r)
  res = ode(Pinitial,tm, hfunc, gps)
  colnames(res)=c("time","carbon")
  res=as.data.frame(res)
  return(carbon=res$carbon)
}

# apply this function to all harvest values
res = harvestr %>% map_dfc(~getcarbon( Pinitial=Pinitial, tm=tm, K=100, r=0.2, hfun=dharvest, harv=.x))
```


```{r harvest}
# rows are time, columns are carbon for each harvest scenario
colnames(res)=harvestr #name columns 
res=as.data.frame(res) #dataframe 
res$time=tm
```


```{r harvest}
# put in to a form where we can plot
resl = gather(res, key="harvestr",value="carbon", -time)
ggplot(resl,aes(time, carbon, col=harvestr)) + 
  geom_line() + 
  theme_classic()
```


```{r harvest}
# notice that stable forest value changes with harvest rates

# notes that some forests are not stable - (or stablity is zero)

# see this at the beginning - plot the first 10 years
ggplot(subset(resl, time < 10),aes(time, carbon, col=harvestr)) + 
  geom_line() + 
  theme_classic()

#plot only the first 10 years to better see how higher harvest rates are too high and keep the forest at 0 

```

# Fixed harvest rate 

Try different harvest rates...notice how carbon growth changes...what does stability mean - can you find parameter sets that lead to a stable non-zero forest

We could do some sensitivity analysis to see how harvest rate, and growth rate interact to control where the forest ends up after 10 years and 50 years (short and long planning horizons)

We will use our compute metrics and wrapper to make this easy

```{r harvestsend}
# fixed amount per year
source(here("R", "dharvest.R"))
```

We can do a normal Sobel analysis 
```{r harvestsend}
# lets assume a uniform distribution of harvest rates
# and of normal growth rates

#we will only vary r and the harvest rate 

np=200

r = rnorm(mean=0.3, sd=0.05, n=np) #0.3 growth rate with some uncertainty around it 
harv = runif(min=0.0, max=0.4, n=np)
X1 = cbind.data.frame(r=r, harv=harv)
```


```{r harvestsend}
# repeat to get our second set of samples
r = rnorm(mean=0.3, sd=0.05, n=np)
harv = runif(min=0.0, max=0.4, n=np)
X2 = cbind.data.frame(r=r, harv=harv)
```


```{r harvestsend}
# create our sobel object and get sets ofparameters for running the model

sens_forest = sobolSalt(model = NULL,X1, X2, nboot = 300)
colnames(sens_forest$X)= c("r","harv")
```

```{r harvestsend}
# do a quick test
# try it out
tm = seq(from=1, to=50)
Pinitial = 1
parms = list(r=sens_forest$X[1,1], harv=sens_forest$X[1,2], K=100)
res = ode(y=Pinitial,times=tm, func=dharvest, parms=parms)
res=as.data.frame(res)
colnames(res)=c("time","C")

# compute our two metrics of interest - harvest after 10 and 50 years
compute_metrics = function(res) {
 C50 = res[50]
 C10 = res[10]
return(list(C50=C50, C10=C10))}
```


```{r harvestsend}
# use a wrapper function to just return the carbon trajectories
p_wrapper = function(r,harv, K, Pinitial, simtimes, func) {
    parms = list(r=r, K=K, harv=harv) #same value for K 
    result = ode(y=Pinitial, times=simtimes, func=func, parms=parms) 
    result=as.data.frame(result)
   colnames(result)=c("time","C")
  # get metrics
  metrics=compute_metrics(result$C)
  return(metrics)
}

# notice how we added in K, a parameter that we are NOT varying

#Sobel only knows what r and the harvest rate are 
```


```{r harvestsend}
# try it out
tm = seq(from=1, to=50)
Pinitial = 1


allresults = as.data.frame(sens_forest$X) %>% pmap(p_wrapper, K=100, Pinitial=Pinitial, simtimes=tm, func=dharvest)

# extract out results from pmap into a data frame
allres = allresults %>% map_dfr(`[`,c("C10","C50"))

tmp = allres %>% gather(key="metric", value="value")
ggplot(tmp, aes(metric, value, col=metric))+geom_boxplot()

#it makes sense that c50 (year 50) is larger than c10 (year 10)
#c50 also has a larger box = more uncertainty over time 
```

Look at the response of c50 to the harvest rate 
- we can view the tradeoffs between the growth rate and harvest rate 
```{r harvestsend}
# link with parameters to see how parameters together
# impact something we care aboue C50

allresp = cbind.data.frame(sens_forest$X, allres)
ggplot(allresp, aes(harv, C50, col=r))+geom_point()

# notice how we can see the impact of harvesting - and how growth rates reduce the sustainable harvest
```

# May 12: Stability 

```{r}
# given some forest characteristics - lets look at derivatives under different harvest rates
lowHrate = 0.015
gps = list(harv=lowHrate, K=100, r=0.05)

# look at the derivative over a range of forest sizes

findstable = data.frame(Ccurr=seq(from=1, to=100, by=5))
# notice use of Time=NULL, and map to compute derivative
# for different values of forest biomass
findstable$dervHlow= unlist(findstable$Ccurr %>% map(~dharvest(parms=gps, Time=NULL, P=.x) ))
                                                  
ggplot(findstable, aes(Ccurr, dervHlow))+geom_point()+geom_hline(yintercept = 0, col="red")+
  labs(y="Derivative\n (Rate of change of forest carbon) (kg/C/year)", x="Current Forest Carbon Stock (kgC)")

```

```{r}
# Populations will be stable when derivative is zero!

# look at a different harvest rate
midHrate=0.02
gps = list(harv=midHrate, K=100, r=0.05)
findstable$dervHmid= unlist(findstable$Ccurr %>% map(~dharvest(parms=gps, Time=NULL, P=.x) ))
 
# try high rate
highHrate=0.05
gps = list(harv=highHrate, K=100, r=0.05)
findstable$dervHhigh= unlist(findstable$Ccurr %>% map(~dharvest(parms=gps, Time=NULL, P=.x) ))
 

# plot them all together
tmp = gather(findstable, key="HarvestRate", value="value", -Ccurr)
ggplot(tmp, aes(Ccurr, value, color=HarvestRate))+geom_point()+geom_hline(yintercept = 0, col="black")+
  labs(x="Forest Biomass (kgC)", y="Forest Growth Rate (kgC/year)")

# notice how with higher harvest rates the stable population will be lower
```


```{r}
tm = seq(from=1, to=500)
gps = list(harv=lowHrate, K=100, r=0.05)
Pinitial=50
res = ode(Pinitial,tm, dharvest, gps)
colnames(res)=c("time","carbon")
ggplot(as.data.frame(res), aes(time, carbon))+geom_point()+
  labs(y="Forest Biomass (kgC)", x="Year", title="low harvest rate")

```

```{r}
gps = list(harv=highHrate, K=100, r=0.05)
Pinitial=50
res = ode(Pinitial,tm, dharvest, gps)
colnames(res)=c("time","carbon")
ggplot(as.data.frame(res), aes(time, carbon))+geom_point()+
 labs(y="Forest Biomass (kgC)", x="Year", title="High Harvest Rate")
```

