---
title: "class_exercise"
author: "Mia Forsline"
date: "4/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Another example: Okuns Law (conceptual, abstract model) - try implementing this as a function

Graph of US quarterly data (not annualized) from 1947 through 2002 estimates a form of the difference version of

**Okun’s law:**

%Change GDP = 0.856 - 1.827(Change Unemployment Rate)

R^2 of .504.

Differences from other results are partly due to the use of quarterly data http://en.wikipedia.org/wiki/Okun%27s_law

Write the function in R!

```{r}
okuns_law <- function(perc_unemployment, intercept = 0.856, slope = 1.827){
  perc_change_gdp = intercept - (slope * perc_unemployment)
  return(perc_change_gdp)
}
```

what is change in GDP for unemployment rate change of 3%

```{r}
okuns_law(perc_unemployment = 3)
```


what would the change in GDP be for 3% change in unemployment rate if the slope was 10% higher

```{r}
okuns_law(perc_unemployment = 3, slope = 1.827 * 1.1)
```

